---
title: "HW1"
output:
  github_document
---


```{r setup, include=FALSE}
rm(list=ls())
library(rmarkdown)
knitr::opts_chunk$set(echo = FALSE)
```

```{r, include=FALSE}
library(ggplot2)
library(tidyr)
library(dplyr)
setwd("~/Desktop/statistical learning/code/data")
ABIA <- read.csv("ABIA.csv")
```



## Flights at ABIA

In this question, we explore the best time to fly to and depart from Austin-Bergstrom Interational Airport in order to minimize delay.

&NewLine;

</br>

#####data-preprocessing

 - new variable 'type': label the flight type as 'Departure' from or 'Arrival' at Austin.
 - new variable 'delaytime': use the departure delay 'DepDelay' as the delay time for departure flights, and arrival delay 'ArrDelay' for arrivals.
 - modify 'delaytime': normalize the delay time to 0 for early departures and arrivals (delay time being negative), assuming that people only care about delays.
  
&NewLine;


</br>

###Seasonality of delays - best time of the year to travel from & to Austin?

&NewLine;

</br>

#####data-processing for seasonality

 - new variable 'date': format the month and date into one variable.
 - calculate the average delay time of departure & arrival flights seperately, across all dates of the year.


```{r, include=FALSE}
#create variable: delaytype = departure / arrival, and their corresponding delay time
ABIA$delaytime <- ifelse(ABIA$Origin == "AUS", ABIA$DepDelay, ABIA$ArrDelay)
ABIA$delaytype <- ifelse(ABIA$Origin == "AUS", "Departure", "Arrival")

#normalize delay time to be 0 or positive
ABIA$delaytime <- ifelse(ABIA$delaytime < 0, 0, ABIA$delaytime)

# create variable: date (yy-mm-dd)
ABIA$date <- with(ABIA, ISOdate(year = 2008, Month, DayofMonth))
```

#####plotting

 - plot the average delay time for two types of flights across all dates.
 - add the trend line for departure and arrival flights (method: lasso)

  
```{r}
season <- ABIA %>%
  select (date, delaytime, delaytype) %>%
  group_by (delaytype, date) %>%
  summarise(avg_delay = mean(delaytime, na.rm= TRUE)) %>%
  ungroup() %>%
  arrange(-avg_delay)
```

&NewLine;

</br>

```{r, warning=FALSE, message=FALSE}
g0 <- ggplot(data=season, aes(date, avg_delay))+
  labs(
    x = "Date",
    y = "Average Delay (min)",
    title = "Seasonality of delay time",
    caption = "Data: commercial flights in 2008 departed from or landed at Austin-Bergstrom Interational Airport"
  )
 
g1 <- g0 +
  geom_point(aes(color=delaytype)) +
  ylim(0,50)

g1 + 
  geom_smooth(mapping = aes(color = delaytype))+
  #geom_smooth(color="blue", linetype="dashed", se=FALSE)+
  labs(
    colour = "Type",
    subtitle = "    Arrival is expected to have longer delay time than departure across all months"
  )
```

&NewLine;
&NewLine;


</br>
</br>

 - plot the trend line for all flights from and to Austin and zoom in to see the details.
  
&NewLine;


</br>

```{r, warning=FALSE, message=FALSE}
g1 + geom_smooth(color="blue")+
  labs(
    colour = "Type",
    subtitle = "    average delay time of departures and arrivals across all months"
  )
```

&NewLine;

</br>

  
&NewLine;

</br>

```{r, warning=FALSE, message=FALSE}
g0 + 
  xlab("Date") + ylab("Average Delay (min)") + geom_smooth(color = "Blue") +
  labs(subtitle = "    Fall is the best season of the year to minimize delays; peak in delays during December")

```

&NewLine;

</br>

We see that throughout the year, arrival flights at Austin are expected have longer delays than departure flights from Austin. Overall, fall (from September to November) is the best season of the year to minimize delays whereas December is the worst month of the year.


&NewLine;

</br>

###Delays in a day - best hour of the day to travel from & to Austin?

&NewLine;

</br>

##### data-processing for delays in a day

 - new variable 'time': use the scheduled departure time 'CRSDepTime' as the time for departure flights, and scheduled arrival time 'CRSArrTime' for arrivals.

```{r}
ABIA$time <- ifelse(ABIA$Origin == "AUS", ABIA$CRSDepTime, ABIA$CRSArrTime)
ABIA$hour=floor(ABIA$time/100)+round((ABIA$time/100-floor(ABIA$time/100))/60*100)
ABIA$hour <- ifelse(ABIA$hour == 24, 0, ABIA$hour)
```
 - new variable 'hour': round the time to the nearest hour.
 - calculate the average delay time of departure & arrival flights seperately, across all hours of the day.
```{r, include=TRUE, echo=TRUE}
summary(ABIA$time)
```
  
  
```{r}
hour <- ABIA %>%
  select (hour, delaytime, delaytype) %>%
  filter (delaytime >= 0) %>%
  group_by (delaytype, hour) %>%
  summarise(avg_delay = mean(delaytime, na.rm= TRUE)) %>%
  ungroup() %>%
  arrange(-avg_delay)%>%
  na.omit()
```

&NewLine;

</br>

#####plotting

  - plot the average delay time for two types of flights across all hours.
  - add the trend line for flights departed from and arrived at Austin (method: lasso).
  
&NewLine;

</br>

```{r, message=FALSE}
ggplot(data=hour, aes(hour, avg_delay)) +
  geom_point(aes(color=delaytype))+ geom_smooth() + 
  xlim(0,24) +
  labs(
    x = "Hour of the day",
    y = "Average Delay (min)",
    color = "Type",
    title = "Hourly delay time",
    subtitle = "    Morning (5 am - 10 am) might be the best time of the day to minimize delays",
    caption = "Data: commercial flights in 2008 departed from or landed at Austin-Bergstrom Interational Airport"
  )

```

&NewLine;

</br>

We see that average delay time increases from 9:00 to 24:00, and reaches peak at midnight. Morning (5 am - 10 am) might be the best time of the day to minimize delays.


Overall, the best way to minimize delay in Austin-Bergstrom Interational Airport is to depart from Austin in the morning of October!

&NewLine;
&NewLine;


</br>
</br>




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Regression vs KNN

```{r, include=FALSE}
library(tidyverse)
library(FNN)
library(mosaic)
library(ggplot2)
```

Extract two trims(350 and 65 AMG) from all trim levels.
```{r, include=FALSE}
# read data 
sclass <- read.csv("~/Documents/GitHub/ECO395M/data/sclass.csv")

# two category
sclass350 = subset(sclass, trim == '350')

sclass65AMG = subset(sclass, trim == '65 AMG')

```

1.Split the data into training and testing set.

```{r,echo=FALSE}
# training and testing data
N1 = nrow(sclass350)
N1_train = floor(0.8*N1)
N1_test = N1 - N1_train

N2 = nrow(sclass65AMG)
N2_train = floor(0.8*N2)
N2_test = N2 - N2_train
# sample a set of data in train data
train_1 = sample.int(N1, N1_train, replace=FALSE)
train_2 = sample.int(N2, N2_train, replace=FALSE)

```

```{r}
# define training and testing data
D1_train = sclass350[train_1,]
D1_test = sclass350[-train_1,]

D2_train = sclass65AMG[train_2,]
D2_test = sclass65AMG[-train_2,]

D1_test = arrange(D1_test, mileage)
D2_test = arrange(D2_test, mileage)

```
```{r, include=FALSE}
# define name tag
X1_train = select(D1_train, mileage)
y1_train = select(D1_train, price)
X1_test = select(D1_test, mileage)
y1_test = select(D1_test, price)

X2_train = select(D2_train, mileage)
y2_train = select(D2_train, price)
X2_test = select(D2_test, mileage)
y2_test = select(D2_test, price)

```

2.Run K-nearest-neighbors, for many different values of K, starting at K=3. For each value of K, fit the model to the training set and make predictions on your test set.

3.Calculate the out-of-sample root mean-squared error (RMSE) for each value of K.
```{r, include=FALSE}
# define rmse function
rmse = function(y, ypred) {
  sqrt(mean(data.matrix((y-ypred)^2)))
}
```

```{r}
rmse1.matrix= matrix(NA,nrow = 328, ncol = 2)
for (i in 3:330) {
  knn=knn.reg(train = X1_train, test = X1_test, y= y1_train, k=i)
  ypred_knn = knn$pred
  rmse(y1_test, ypred_knn)
  rmse1.matrix[i-2,2]=rmse(y1_test, ypred_knn)
  rmse1.matrix[i-2,1]=i
}

rmse2.matrix= matrix(NA,nrow = 228, ncol = 2)
for (i in 3:230) {
  knn=knn.reg(train = X2_train, test = X2_test, y= y2_train, k=i)
  ypred_knn = knn$pred
  rmse(y1_test, ypred_knn)
  rmse2.matrix[i-2,2]=rmse(y2_test, ypred_knn)
  rmse2.matrix[i-2,1]=i
} 
```


Make plot of RMSE versus K for trim 350


```{r,echo=FALSE}
#plot rmse vesus k
plot(rmse1.matrix,type = "l",xlab = "k", ylab = "rmse", main = "350")
```


Make plot of RMSE versus K for trim 65AMG


```{r,echo=FALSE}
#plot rmse vesus k
plot(rmse1.matrix,type = "l",xlab = "k", ylab = "rmse", main = "65AMG")

```

For the optimal value of K, the plot of fitted model for trim 350
The optimal value of K is below:
```{r,include= FALSE}
k1.matrix= matrix(NA, nrow =100, ncol= 1)
k2.matrix= matrix(NA, nrow =100, ncol= 1)
# all loop
for (n in 1:100) {
  
# sample a set of data in train data
train_1 = sample.int(N1, N1_train, replace=FALSE)
train_2 = sample.int(N2, N2_train, replace=FALSE)

# define data name tag
D1_train = sclass350[train_1,]
D1_test = sclass350[-train_1,]

D2_train = sclass65AMG[train_2,]
D2_test = sclass65AMG[-train_2,]

D1_test = arrange(D1_test, mileage)
D2_test = arrange(D2_test, mileage)

X1_train = select(D1_train, mileage)
y1_train = select(D1_train, price)
X1_test = select(D1_test, mileage)
y1_test = select(D1_test, price)

X2_train = select(D2_train, mileage)
y2_train = select(D2_train, price)
X2_test = select(D2_test, mileage)
y2_test = select(D2_test, price)


# define KNN loop
rmse1.matrix= matrix(NA,nrow = 328, ncol = 1)
for (i in 3:330) {
 knn=knn.reg(train = X1_train, test = X1_test, y= y1_train, k=i)
 ypred_knn = knn$pred
 rmse(y1_test, ypred_knn)
 rmse1.matrix[i-2,1]=rmse(y1_test, ypred_knn)
}

rmse2.matrix= matrix(NA,nrow = 228, ncol = 1)
for (i in 3:230) {
  knn=knn.reg(train = X2_train, test = X2_test, y= y2_train, k=i)
  ypred_knn = knn$pred
  rmse(y1_test, ypred_knn)
  rmse2.matrix[i-2,1]=rmse(y2_test, ypred_knn)
}

# find minimum rmse
apply(rmse1.matrix, 2, min)
k1.matrix[n,1]=which.min(rmse1.matrix)+2

apply(rmse2.matrix, 2, min)
k2.matrix[n,1]=which.min(rmse2.matrix)+2

}

```
```{r,echo= FALSE}
#use mean to find minimum k1
k1=mean(k1.matrix)
k1=floor(k1)
k1

```
The corresponding RMSE is below:
```{r,echo=FALSE}
knnk1 = knn.reg(train = X1_train, test = X1_test, y = y1_train, k=k1)
ypred_knnk1 = knnk1$pred
rmse(y1_test, ypred_knnk1)
```

The plot of fitted model for trim 350


```{r,echo=FALSE}
D1_test$ypred_knnk1 = ypred_knnk1
p1_test = ggplot(data = D1_test) + 
  geom_point(mapping = aes(x = mileage, y = price), color='lightgrey') + 
  theme_bw(base_size=18) + 
  ylim(6000, 120000)
p1_test
p1_test + geom_point(aes(x = mileage, y = ypred_knnk1), color='red')
```


For the optimal value of K, the plot of fitted model for trim 65AMG
The optimal value of K is below:
```{r,echo=FALSE}
k2=mean(k2.matrix)
k2=floor(k2)
k2
```

The corresponding RMSE is below:

```{r,echo=FALSE}
knnk2 = knn.reg(train = X2_train, test = X2_test, y = y2_train, k=k2)
ypred_knnk2 = knnk2$pred
rmse(y2_test, ypred_knnk2)

```

The plot of fitted model for trim 65AMG


```{r,echo=FALSE}
D2_test$ypred_knnk2 = ypred_knnk2
p2_test = ggplot(data = D2_test) + 
  geom_point(mapping = aes(x = mileage, y = price), color='lightgrey') + 
  theme_bw(base_size=18) + 
  ylim(18000, 250000)
p2_test
p2_test + geom_point(aes(x = mileage, y = ypred_knnk2), color='red')

```



##Which trim yields a larger optimal value of K? Why do you think this is?
Trim 350 yeilds larger optimal value of K. Because the sample size for trim 350 is larger than trim 65AMG. Then I guess the larger the sample size is, the larger the optimal K will be.


